\documentclass[smaller,dvipsnames,ratio=169]{beamer}

\usetheme[numbering=fraction,%
          block=fill,%
          sectionpage=progressbar,%
          subsectionpage=progressbar,%
  ]{metropolis} % Use metropolis theme
\setbeamercovered{invisible}

\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{tikz}
  \usetikzlibrary{arrows} % required in the preamble
\usepackage{listings}
\usepackage{todonotes}

\title{SAT Solver Front-End}
\subtitle{Team White}
\author{Filippo~De~Bortoli \and Aneta~Koleva \and Lorenz~Leutgeb}
\institute{Free University of Bozen-Bolzano\\[2mm] \texttt{\{\href{mailto:filippo.debortoli@stud-inf.unibz.it}{filippo.debortoli},\href{mailto:aneta.koleva@stud-inf.unibz.it}{aneta.koleva},\href{mailto:lorenz.leutgeb@stud-inf.unibz.it}{lorenz.leutgeb}\}\newline @stud-inf.unibz.it}}
\date{20-04-2018}

\begin{document}

  \maketitle

  \begin{frame}{Outline}
    \tableofcontents
  \end{frame}

  \section{Introduction}

  \begin{frame}{Task}
	\begin{enumerate}
		\item Given a grammar for a Boolean logic language with standard connectives and quantifiers over finite domains.
		\item Develop an optimizing compiler that translates input formulae into CNF.
		\item Interface with a SAT solver to compute models for input formulae.
	\end{enumerate}
  \end{frame}

  \begin{frame}{Tools of the Trade}
	\begin{enumerate}
		\item Pragmatic choices to avoid any confusions
		\item Keeping it DRY (Don't Repeat Yourself)
		\item Avoid reinvention
	\end{enumerate}

  	\begin{center}
  	\begin{tabular}{lll}
  		{\bfseries Tool/Library} & {\bfseries Version} & {\bfseries Purpose} \\
  		ANTLR & 4.7.1 & Parser Generator \\
  		Java & 10 & Runtime \\
  		SAT4J & 2.3.5 & Solver Components \\
  		Guava & & Utilities and Data Structures \\
  		JCommander & & Commandline Parsing \\
  		JUnit & 5.1 & Testing \\
  		LOG4J & & Logging \\
  		Gradle & 4.6 & Dependency and Build Management \\
  	\end{tabular}
  	\end{center}
  \end{frame}

  % \section{Front-end architecture}

  \begin{frame}{Big Picture}
    % Panorama on the whole front-end.
    % Here, names only. Descriptions in next slides.
    
    \begin{tabular}{lll}
1 & Parser & $(\exists x \in \{a, b\}) \neg (\exists y \in \{c, d\}) (p(x) \land p(y))$ \\
2 &Compilation Passes & \\
2a & ~ Normalization & (to only have $\land, \lor, \neg$)\\
2b & ~ NNF &$ (\exists x {\in} \{a, b\}) (\forall y {\in} \{c, d\}) (\neg p(x) \lor \neg p(y))$ \\
2b & ~ Standardization &$ (\exists x_1 {\in} \{a, b\}) (\forall x_2 {\in} \{c, d\}) (\neg p(x_1) \lor \neg p(x_2))$ \\
2c & ~ Push and Reorder Qs&$ (\forall x_2 {\in} \{c, d\}) (\exists x_1 {\in} \{a, b\}) (\neg p(x_1) \lor \neg p(x_2))$ \\
3 & Ground &$ (\neg p(a) \lor \neg p(c) \lor \neg p(b)) \land \dots$ \\
4 & Tseitin Transformation \\
5 & CNF Minimization \\
5a & ~ Unit Propagation \\
5b & ~ Hole Elimination \\
6 & DIMACS Generation \\
    \end{tabular}
  \end{frame}

  % Add one slide per each relevant class.

  \section{Algorithms and Design Choices}

  \begin{frame}{Parser}
    \begin{enumerate}
      \item We don't always need to parse languages\dots \newline But when we do, we use a parser generator (like ANTLR).
      \item Straightforward parser generation as build step with Gradle: \texttt{gradle~generateGrammarSource}
      \item At runtime use ANTLR API and consume parse tree.
      \item Map parse tree into our custom AST representation which implements compilation passes.
    \end{enumerate}
  \end{frame}

  \begin{frame}{Clause Accumulator}
    \begin{itemize}
      \item \alert{Bookkeeping} during Tseitin Transformation
      \item Clause storage
      \item Which proposition maps to which high level expression?
      \item Pass on clauses to a 2WL container (borrowed from SAT4J)
      \item \alert{Minimize} size of transformed CNF formula using Unit Propagation and Hole Elimination.
    \end{itemize}
  \end{frame}
  
  \begin{frame}
    \frametitle{Pushing and Reordering Quantifiers}
    \begin{itemize}
      \item Quantifiers are pushed \alert{inwards} \textcolor{gray}{to contain size of grounded formula}.
      \item Consecutive quantifiers should be reordered if:
      \begin{itemize}
        \item variables are \alert{unrelated}
        \item domains are \alert{independent}
        \item outer quantifier is existential, inner is universal
      \end{itemize}
      \item Quantifiers of same type are not reordered (irrelevant for output size).
      \item Reordering aims at \alert{minimization} of size of grounded formula.
    \end{itemize}
  \end{frame}

  \begin{frame}{CNF Conversion}
    \begin{itemize}
      \item Receives a ground formula in NNF
      \item \alert{Tseitin Transformation} is applied\ldots
      \item \ldots{}using only \alert{implications} since NNF ensures positive context
      % \item Use implications only (we always are in a positive context thanks to NNF).
      \item Structural \alert{Sharing} to avoid repeated conversions % hashing expressions and doing memoization
    \end{itemize}
  \end{frame}

  \begin{frame}{Unit Propagation}
    \begin{itemize}
      \item To remove Tseitin-generated propositions
      \item If any variable from original input is removed by propagation, it is added  again by means of a unit clause.
      \item Re-use components of SAT4J (2WL, partial Assignment, \texttt{UnitPropagationListener})
      \item \alert{Hole Elimination} for a sparse DIMACS encoding
    \end{itemize}
  \end{frame}

  \begin{frame}[fragile]{Extensions}
    Extend syntax and semantics of expressions to accomodate
    \begin{itemize}
      \item Term (in)\alert{equality}
      \item \alert{Comparison} of integer expressions
    \end{itemize}
    \begin{verbatim}
expression
: ...

| left=intExpression
    op=(GT|LT|GE|LE|NE|EQ)
  right=intExpression
# arithmeticAtom

| left=term op=(NE|EQ) right=term
# equalityAtom

;
    \end{verbatim}
\end{frame}

  \section{Evaluation}

  \begin{frame}
    \frametitle{The Test Suite}
    Powered by \alert{JUnit}.

    Instance types:
    \begin{itemize}
      \item 22 propositional formulas \textcolor{gray}{connectives, known equivalences}
      \item 24 non-ground formulas \textcolor{gray}{integers with related built-ins, characters}
      \item 4 complex encodings \textcolor{gray}{Sudoku, nested quantifiers optimization}
    \end{itemize}
    Asserted results:
    \begin{itemize}
      \item \alert{Size} of generated CNF \textcolor{gray}{\# clauses and \# variables}
      \item Number of \alert{models}
    \end{itemize}
    Execution \alert{time} is measured.
  \end{frame}

  \section{Conclusion}

  \begin{frame}{Possible Improvements}
  	\begin{itemize}
  		\item Merging compiling passes (currently separated for clarity)
  		\item Pass over pre-filled 2WL data structure to solver.
  		\item String as terms (access length, define concatenation)
  		\item Other solvers as back-ends.
  		\item Syntax Highlighting in REPL.
  		\item Refactoring or reimplementing from scratch.
  		\item Use a faster runtime than Java (go native)
  		\item Generate CNF and propagate in parallel.
   		\item Further optimization that we have missed.
  	\end{itemize}
  \end{frame}

  \begin{frame}{Conclusion}
  	\begin{itemize}  	
  		\item Working Implementation
  		\item Some challenges along the way, but nothing impossible.
  		\item Deliberately not optimized for execution speed, we're still happy.
  		\item Some thought went into the optimizing passes (also trial and error).
  	\end{itemize}
  \end{frame}

  \begin{frame}[standout]
    Thank you!
  \end{frame}

\end{document}